---
title: 'Lab Week 5: Time series and a map!'
author: "Lydia Bleifuss"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Attach packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data","renewables_cons_prod.csv")) %>% 
  clean_names()
```


- Make description all lowercase
- Only keep observations where "description" variable contains "consumpiton"
- Remove any observations where "description" variable contains "total" 

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>%  #use mutate to over-right
  filter(str_detect(description, pattern = "consumption")) %>% #now look for a pattern "consumption" and if isn't there get rid of that row (string detect is a true false logical function (if tru then it stays, if false then it is taken out))
  filter(!str_detect(description, pattern = "total"))#because we now want to exclude, or look for false use !
```


### Convert 'yyymm' column to a date AND then help it know that it's a time frame 
```{r}
#Now we want to work with time component!
#Lubridate is great and makes nice assumptions but you should always check (like month 13! It just didn't make dates for those because that's rediculous)

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>%  #give it the structure of the colum and formae
  mutate(month_sep = yearmonth(yr_mo_day)) %>% #now converting to SIBBLE format because FEAST AND FORMAT wants it with SIBBLE, month_sep is stored as a DATE and YEARMON which is great for time series stuff later
  mutate(value = as.numeric(value)) %>%  #value was read in as a character
  drop_na(month_sep, value) #where month_sep OR value has an NA, get rid of that row

#Make a version where I have the month & year in separate colums: 
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% #label = true means that the new month is a class of ordered factor that has the month abbreviations
  mutate(year = year(yr_mo_day))
```

### Look at it: 

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value,
                                          group = description)) +
  geom_line(aes(color = description))


renew_gg
```


Updating colors with paleteer palettes: SO FREAKING COOL
```{r}
renew_gg + 
  scale_color_paletteer_d("ggsci::uniform_startrek")#chood d because we are choosing from DISCRETE functions
```

### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)#index is the sibble compatable time variable (in this case month_sep)
```

Let's look at our ts data in a couple different ways:
```{r}
renew_ts %>% autoplot(value) #because we told it above what to use, this does all the other work
renew_ts %>% gg_subseries(value) #see the different sources broken up by month across all years, graphing exploration super fast!

#renew_ts %>% gg_season(value) #seasonally, but this breaks a lot so let's make this another way:

season_graph <- ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
```


### Just look at hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value) #blue lines are monthly values! 
#hydro_ts %>%  gg_season(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))
  
```


### What if I want quarterly average consumption for hydro? 
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% #group as a function of yearquarter based on (.) the groups that exist 
  summarize(ave_consumption = mean(value))

head(hydro_quarterly)
```


### Decompose that hydro_ts
```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5))) #model value as a function of season

components(dcmp) %>% autoplot() #+ #top is acutaly data, then moving average tren, then seasonal extracted, then fourth graph is anything is left over 
  #scale_y_continuous(limits = c(150,350)) #you can force scales like you would in gg plot (not always helpful)
```

```{r}
hist(components(dcmp)$remainder)
```

Now look at the ACF:
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

#x axis lab[1M], means that it is showing month by month and that every 12 months are most coorelated 
```


###DANGER DANGER DANGER (don't use this unless you know your shit)
```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ETS(value)
  ) %>% 
  fabletools::forecast(h = "4 years") #how long into the future since your last point you want to forcast

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))#want to tack on existing data and can choose time range

```

### Make a world map!

#customizing use tmap, quick and dirty look use mapview

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

